Continous time Markov chain models; \\ extension, simulation and visualisation

### 2020-12-11 chapters/chapter03.tex
B-W figures for appendices
fixing floats
implementing comments
last minute modifications

\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{RacRABV_cont_vis}
\caption{
{ \footnotesize 
{\bf Screenshot of SPREAD.}
This example visualizes the phylogeographic history of rabies among raccoons along the Eastern US seaboard under a continuous diffusion model.  Such visualizations allow users to quickly inspect key evolutionary changes in their geographic context.  Further generation of KML output enables interactive exploration in the time-dimension as well in freely available virtual globe software, such as {\sc Google Earth} (on right).
}% END: footnotesize 
}
\label{fig:screenshot01}
\end{figure}

@@

The advent of powerful and flexible Geographic Information Systems (GIS) has fostered an increasing interest in incorporating geographical information into molecular phylogenetic methods.
Spatial phylogenetic projections in a cartographic background play an important role in these developments \citep{Kidd2006, Parks2009}, but most applications remain limited to mapping phylogenetic tip taxa to their geographical coordinates.
Mapping phylogeographic histories of geo-referenced taxa requires a robust statistical estimate of the geographic locations at the ancestral nodes of the tree. 
To obtain such estimates under stochastic process-driven models, we recently proposed a suite of Bayesian inference approaches for the joint reconstruction of evolutionary and geographic history \citep{Lemey2009,Lemey2010, Bloomquist2010}.
Models that accommodate spatial diffusion in discrete and continuous space have been implemented in a flexible Bayesian statistical framework (BEAST, \cite{Drummond2007}, \url{http://beast.bio.ed.ac.uk}) for hypothesis testing based on time-measured evolutionary histories.
These approaches produce statistical distributions of temporal-spatial phylogenies and this has created new challenges for statistical phylogenetics in producing informative and compelling visualizations.
Here, we present software to fully exploit spatio-temporal annotations on phylogenies by providing flexible visual summaries that can be further examined in an interactive manner using GIS or virtual globe software.

@@

SPREAD provides four templates to analyze and visualize different aspects of phylogeographic diffusion, labeled: \textbf{Discrete Tree}, \textbf{Discrete Bayes Factors}, \textbf{Continuous Tree} and \textbf{Time Slicer}.
The discrete and continuous tree templates typically provide posterior summaries of diffusion from a Bayesian analysis along a high-probability tree, such as the maximum clade credibility (MCC) tree of BEAST. 
However, SPREAD is not necessarily limited to this input, as it employs a general, trait-annotated NEXUS format; we provide several tree file examples in the supplementary information for both discrete or continuous annotations. 
SPREAD supports customized visualization of spatially-mapped trees, including branch coloring according to time and branch width manipulation.

>>>>

\begin{figure}[h!]
\centering
\includegraphics[scale=0.25]{RacRABV_cont_vis-color}
\caption{
{ \footnotesize 
{\bf Screenshot of SPREAD.}
This example visualizes the phylogeographic history of rabies among raccoons along the Eastern US seaboard under a continuous diffusion model.  Such visualizations allow users to quickly inspect key evolutionary changes in their geographic context.  Further generation of KML output enables interactive exploration in the time-dimension as well in freely available virtual globe software, such as {\sc Google Earth} (on right).
}% END: footnotesize 
}
\label{fig:screenshot01}
\end{figure}

@@

The advent of powerful and flexible Geographic Information Systems (GIS) has fostered an increasing interest in incorporating geographical information into molecular phylogenetic methods.
Spatial phylogenetic projections in a cartographic background play an important role in these developments \citep{Kidd2006, Parks2009}, but most applications remain limited to mapping phylogenetic tip taxa to their geographical coordinates.
Mapping phylogeographic histories of geo-referenced taxa requires a robust statistical estimate of the geographic locations at the ancestral nodes of the tree. 
To obtain such estimates under stochastic process-driven models, we recently proposed a suite of Bayesian inference approaches for the joint reconstruction of evolutionary and geographic history \citep{Lemey2009,Lemey2010, Bloomquist2010}.
Models that accommodate spatial diffusion in discrete and continuous space have been implemented in a flexible Bayesian statistical framework (BEAST, \cite{Drummond2007}, \url{http://beast.bio.ed.ac.uk}) for hypothesis testing based on time-measured evolutionary histories.
These approaches produce statistical distributions of temporal-spatial phylogenies and this has created new challenges for statistical phylogenetics in producing informative and compelling visualizations.
Here, we present software to fully exploit spatio-temporal annotations on phylogenies by providing flexible visual summaries that can be further examined in an interactive manner using GIS or virtual globe software.
A tutorial that explains how to use the software is included as Appendix~\ref{app:spread_tuto}.

@@

SPREAD provides four templates to analyze and visualize different aspects of phylogeographic diffusion, labeled: \textbf{Discrete Tree}, \textbf{Discrete Bayes Factors}, \textbf{Continuous Tree} and \textbf{Time Slicer}.
The discrete and continuous tree templates typically provide posterior summaries of diffusion from a Bayesian analysis along a \newline 
high-probability tree, such as the maximum clade credibility (MCC) tree of BEAST. 
However, SPREAD is not necessarily limited to this input, as it employs a general, trait-annotated NEXUS format; we provide several tree file examples in the supplementary information for both discrete or continuous annotations. 
SPREAD supports customized visualization of spatially-mapped trees, including branch coloring according to time and branch width manipulation.
